---
title: "Geospatial Risk Prediction"
author: "Minwook Kang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cerulean
    highlight: monochrome
    css: custom.css
---

1. A map of your outcome of interest in point form, with some description of what, when, and why you think selection bias may be an issue.

2. A map of your outcome joined to the fishnet.

3. A small multiple map of your risk factors in the fishnet (counts, distance and/or other feature engineering approaches).

4. Local Moran’s I-related small multiple map of your outcome (see 5.4.1)

5. A small multiple scatterplot with correlations.

6. A histogram of your dependent variable.

7. A small multiple map of model errors by random k-fold and spatial cross validation.

8. A table of MAE and standard deviation MAE by regression.

9. A table of raw errors by race context for a random k-fold vs. spatial cross validation regression.

10. The map comparing kernel density to risk predictions for the next year’s crime.

11. The bar plot making this comparison.

12. Two paragraphs on why or why not you would recommend your algorithm be put into production.

<style>
div.blue pre.r { background-color:#eff3ff; }
</style>




## Ⅰ. Setup 

<div class = "blue">
```{r message=FALSE, warning=FALSE}

# Rmarkdown global setting
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(fig.align = 'center')

#----------------------------------------------------------------------------------------------------------  

# Import libraries
library(tidyverse)
library(tidycensus)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)# plot correlation plot
library(corrplot)
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots
library(knitr)
library(rmarkdown)
library(RSocrata)
library(viridis)
library(ggplot2)
library(stargazer)
library(XML)
library(data.table)
library(ggpmisc)
library(patchwork)
library(spatstat)
library(raster)
library(classInt)   # for KDE and ML risk class intervals

#----------------------------------------------------------------------------------------------------------  

# Temp
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

# Etc
options(scipen=999)
options(tigris_class = "sf")

#----------------------------------------------------------------------------------------------------------  

# functions

st_c    <- st_coordinates
st_coid <- st_centroid

mapThememin <- function(base_size = 10, title_size = 12, small_size = 8) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = title_size, colour = "black", hjust = 0.5),
    plot.subtitle=element_text(size = base_size, colour = "black", hjust = 0.5, face="italic"),
    plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.text.x = element_text(size = base_size),
    strip.text.y = element_text(size = base_size),
    strip.background = element_rect(colour="transparent", fill="transparent"),
    legend.title = element_text(size = small_size),
    legend.text = element_text(size = small_size),
    legend.key.size = unit(0.4, "cm"))
}


corTheme <- function(base_size = 10, title_size = 12, small_size = 8){
  theme(axis.text =  element_blank(), 
        axis.ticks = element_blank(), 
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}

corTheme2 <- function(base_size = 10, title_size = 12, small_size = 8){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}

corTheme3 <- function(base_size = 9, title_size = 11, small_size = 7){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}

corTheme4 <- function(base_size = 9, title_size = 11, small_size = 7){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y.right = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}


q5 <- function(variable) {as.factor(ntile(variable, 5))}

q <- function(variable) {as.factor(ntile(variable, 5))}

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                                  c(.01,.2,.4,.6,.8), na.rm=T),
                         digits = 3))
  }
}

qBr2 <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]]*100,0)/100,
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                                  c(.01,.2,.4,.6,.8), na.rm=T),
                         digits = 3))
  }
}

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <- get.knnx(measureTo, measureFrom, k)$nn.dist
  output <- as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  return(output)  
}

#----------------------------------------------------------------------------------------------------------  

# Colors ("https://coolors.co/gradient-palette/a8f368-f9035e?number=7")
bluePalette5 <- c("#eff3ff","#bdd7e7","#6baed6","#3182bd","#08519c")
blue2Palette5 <- c("#08519c","#3182bd","#6baed6","#bdd7e7","#eff3ff")
orangePalette5 <- c("#FFF2E8","#FFD6B6","#FEB984","#FE9D51","#FD801F")
orange2Palette5 <- c("#FFDFD0","#FFB89F","#FF926E","#FF6B3D","#FF440C")
greyPalette5 <- c("#f7f7f7","#cccccc","#969696","#636363","#252525")
greenPalette5 <- c("#edf8e9","#bae4b3","#74c476","#31a354","#006d2c")
purplePalette5 <- c("#f2f0f7","#cbc9e2","#9e9ac8","#756bb1","#54278f")

#----------------------------------------------------------------------------------------------------------  

# LoadAPI(Min's key)
census_api_key("4bbe4bead4e5817f6a6b79e62c5bea69e77f1887", overwrite = TRUE)


```
</div>



## Ⅱ. Data

<div class = "blue">
```{r message=FALSE, warning=FALSE, include = FALSE, cache=TRUE}

# Tracts geometry, Census Information

policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271')

neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson")%>%
  st_transform('ESRI:102271')

#----------------------------------------------------------------------------------------------------------  

# update to 2018 data

# some of risk factors data hasn't been updated yet after 2018

Criminals <- read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2018/3i3m-jwuy")

#----------------------------------------------------------------------------------------------------------  

# check criminal distributions 

CriminalsType <- Criminals %>% dplyr::select(Primary.Type) %>% group_by(Primary.Type) %>% summarise(count = n())

CriminalsDescription <- Criminals %>% dplyr::select(Description) %>% group_by(Description) %>% summarise(count = n())

#----------------------------------------------------------------------------------------------------------  
  
# Dependent variable : MOTOR VEHICLE THEFT - AUTOMOBILE

MVTheft <- Criminals %>%
  filter(Primary.Type == "MOTOR VEHICLE THEFT" & Description == "AUTOMOBILE") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),Y = as.numeric(Y)) %>%
  dplyr::select(-Date, -Updated.On) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
  st_transform('ESRI:102271') %>%
  distinct()

#----------------------------------------------------------------------------------------------------------  

# Creating a fishnet grid

fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 500, 
               square = TRUE) %>%
  .[chicagoBoundary] %>%            # fast way to select intersecting polygons
  st_sf() %>%
  mutate(uniqueID = 1:n())

#----------------------------------------------------------------------------------------------------------  

# 6 risk factors

# risk factor 1 of 6 : Abandoned cars

rf1_abandonCars <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva")

rf1 <- rf1_abandonCars %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Cars")

# risk factor 2 of 6 : Abandoned Buildings

rf2_abandonBuildings <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/7nii-7srd") 

rf2 <- rf2_abandonBuildings %>%
    mutate(year = substr(date_service_request_was_received,1,4)) %>%  filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Buildings")

# risk factor 3 of 6 : graffiti

rf3_graffiti <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Graffiti-Removal-Historical/hec5-y4x5")

rf3 <- rf3_graffiti %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    filter(where_is_the_graffiti_located_ %in% c("Front", "Rear", "Side")) %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Graffiti")

# risk factor 4 of 6 : street Lights Out

rf4_streetLightsOut <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Street-Lights-All-Out/zuxi-7xem")

rf4 <- rf4_streetLightsOut %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Street_Lights_Out")
    
    
# risk factor 5 of 6 : sanitation

rf5_sanitation <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Sanitation-Code-Complaints-Hi/me59-5fac")

rf5 <- rf5_sanitation %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Sanitation")
    
    

# risk factor 6 of 6 : liquor Retail

rf6_liquorRetail <- read.socrata("https://data.cityofchicago.org/resource/nrmj-3kcf.json")

rf6 <- rf6_liquorRetail %>%
    filter(business_activity == "Retail Sales of Packaged Liquor") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Liquor_Retail")


```
</div>


## Ⅲ. Process

**1. Points Map of Automobile Theft**

description of what, when, and why you think selection bias may be an issue

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary, fill = greyPalette5[2], color = "transparent", alpha = 0.5) +
  geom_sf(data = MVTheft, colour=orangePalette5[4], size=0.01, show.legend = "point") +
  labs(title= "Automobile Theft",
       subtitle = "Total 8,800 cases",
       caption = "Figure 1-1") +
  mapThememin(),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = greyPalette5[2], color = "transparent", alpha = 0.5) +
  stat_density2d(data = data.frame(st_coordinates(MVTheft)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = orangePalette5[1], high = orangePalette5[5]) +
  scale_alpha(range = c(0.00, 0.2), guide = FALSE) +
  labs(title = "Density of Automobile Theft",
       subtitle = "2018 Chicago",
       caption = "Figure 1-2") +
  mapThememin() + theme(legend.position = "none"))

```
</div>


**2. Aggregate Automobile Theft to the Fishnet**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

crime_net <- 
  dplyr::select(MVTheft) %>% 
  mutate(countMVTheft = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countMVTheft = replace_na(countMVTheft, 0),
         uniqueID = 1:n(),
         cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE)) # cross-validation id approx 100 folds

ggplot() +
  geom_sf(data = crime_net, aes(fill = q5(countMVTheft)), color = "transparent")+
  scale_fill_manual(values = orangePalette5,
                    labels = qBr(crime_net, "countMVTheft"),
                    name = "Automobile\nTheft counts\n(quantiles)",
                    na.translate=FALSE) +
  labs(title = "Aggregate Map of Automobile Theft",
       subtitle = "2018 Chicago",
       caption = "Figure 2-1") +
  mapThememin()


```
</div>


**3. Multiple map of Automobile Theft risk factors in the fishnet**

counts, distance and/or other feature engineering approaches

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


vars_net <- 
  rbind(rf1,rf2,rf3,rf4,rf5,rf6) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend, count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()


vars_net <- vars_net %>%
    mutate(Abandoned_Cars_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf1), k = 3)),
           Abandoned_Buildings_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf2), k = 3)),
           Graffiti_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf3), k = 3)),
           Street_Lights_Out_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf4), k = 3)),
           Sanitation_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf5), k = 3)),
           Liquor_Retail_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf6), k = 3)))

vars_net.nn <- 
  dplyr::select(vars_net, ends_with("nn3")) %>%
    gather(Variable, value, -geometry)

vars <- unique(vars_net.nn$Variable)
mapList <- list()


for(i in vars){
  mapList[[i]] <-
    ggplot() +
    geom_sf(data = filter(vars_net.nn, Variable == i), aes(fill=q5(value)), color = "transparent") +
    scale_fill_manual(values = bluePalette5,
                      labels = qBr(vars_net.nn, "value"),
                      name = "Automobile\nTheft counts\n(quantiles)",
                      na.translate=FALSE) +
    labs(title = i) +
  mapThememin()
}

do.call(grid.arrange,c(mapList, ncol=3, top="Risk Factors by Fishnet"))

for(i in vars){
  mapList[[i]] <-
    ggplot() +
    geom_sf(data = filter(vars_net.nn, Variable == i), aes(fill=value), color = "transparent") +
    scale_fill_gradient(low = bluePalette5[1], high = bluePalette5[5]) +
    labs(title = i) +
  mapThememin()
}

do.call(grid.arrange,c(mapList, ncol=3, top="Risk Factors by Fishnet"))

#----------------------------------------------------------------------------------------------------------

```
</div>


**4. Local Moran’s I-related small multiple map of your outcome**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

## generates warnings from PROJ issues
## {spdep} to make polygon to neighborhoods... 
final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)
## ... and neighborhoods to list of weigths
final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)

# print(final_net.weights, zero.policy=TRUE)

## see ?localmoran
local_morans <- localmoran(final_net$Abandoned_Cars, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

# join local Moran's I results to fishnet
final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(Abandoned_Cars_Count = Abandoned_Cars, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)

## This is just for plotting
vars <- unique(final_net.localMorans$Variable)
varList <- list()

for(i in vars){
  varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme(title_size = 14) + theme(legend.position="bottom")}

do.call(grid.arrange,c(varList, ncol = 4, top = "Local Morans I statistics, Burglary"))

```
</div>


**5. Multiple Scatterplots with correlations**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

correlation.long <-
  st_drop_geometry(final_net) %>%
    dplyr::select(-uniqueID, -cvID, -loopDistance, -name, -District) %>%
    gather(Variable, Value, -countBurglaries)

correlation.cor <-
  correlation.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, countBurglaries, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, countBurglaries)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Burglary count as a function of risk factors") +
  plotTheme()

```
</div>


**6. A histogram of dependent variable**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>



**7. Maps of model errors**

by k-fold and spatial cross validation

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


**8. A table of MAE and standard deviation MAE by regression**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


**9. A table of raw errors by race context**

for a random k-fold vs. spatial cross validation regression.

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


**10. A table of raw errors by race context for a random k-fold vs. spatial cross validation regression.**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


**11. The bar plot making this comparison.**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


**12. The bar plot making this comparison.**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


```
</div>


## Ⅳ. Results




























































