---
title: "Geospatial Risk Prediction"
author: "Minwook Kang & Shengao Yi"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: cerulean
    highlight: monochrome
    css: custom.css
---

1. A map of your outcome of interest in point form, with some description of what, when, and why you think selection bias may be an issue.

2. A map of your outcome joined to the fishnet.

3. A small multiple map of your risk factors in the fishnet (counts, distance and/or other feature engineering approaches).

4. Local Moran’s I-related small multiple map of your outcome (see 5.4.1)

5. A small multiple scatterplot with correlations.

6. A histogram of your dependent variable.

7. A small multiple map of model errors by random k-fold and spatial cross validation.

8. A table of MAE and standard deviation MAE by regression.

9. A table of raw errors by race context for a random k-fold vs. spatial cross validation regression.

10. The map comparing kernel density to risk predictions for the next year’s crime.

11. The bar plot making this comparison.

12. Two paragraphs on why or why not you would recommend your algorithm be put into production.

<style>
div.blue pre.r { background-color:#eff3ff; }
</style>




## Ⅰ. Setup 

<div class = "blue">
```{r message=FALSE, warning=FALSE}

# Rmarkdown global setting
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(fig.align = 'center')

#----------------------------------------------------------------------------------------------------------  

# Import libraries
library(tidyverse)
library(tidycensus)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)# plot correlation plot
library(corrplot)
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots
library(knitr)
library(rmarkdown)
library(RSocrata)
library(viridis)
library(ggplot2)
library(stargazer)
library(XML)
library(data.table)
library(ggpmisc)
library(patchwork)
library(spatstat)
library(raster)
library(classInt)   # for KDE and ML risk class intervals
library(tableHTML)

#----------------------------------------------------------------------------------------------------------  

# Temp
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

# Etc
options(scipen=999)
options(tigris_class = "sf")

#----------------------------------------------------------------------------------------------------------  

# functions

st_c    <- st_coordinates
st_coid <- st_centroid

mapThememin <- function(base_size = 10, title_size = 12, small_size = 8) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = title_size, colour = "black", hjust = 0.5),
    plot.subtitle=element_text(size = base_size, colour = "black", hjust = 0.5, face="italic"),
    plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.background = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    strip.text.x = element_text(size = base_size),
    strip.text.y = element_text(size = base_size),
    strip.background = element_rect(colour="transparent", fill="transparent"),
    legend.title = element_text(size = small_size),
    legend.text = element_text(size = small_size),
    legend.key.size = unit(0.4, "cm"))
}


corTheme <- function(base_size = 10, title_size = 12, small_size = 8){
  theme(axis.text =  element_blank(), 
        axis.ticks = element_blank(), 
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5),
        strip.background = element_rect(colour="transparent", fill="transparent"))
}


corTheme2 <- function(base_size = 10, title_size = 12, small_size = 8){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5),
        strip.background = element_rect(colour="transparent", fill="transparent"))
}

corTheme3 <- function(base_size = 9, title_size = 11, small_size = 7){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}

corTheme4 <- function(base_size = 9, title_size = 11, small_size = 7){
  theme(axis.text =  element_text(size = small_size),
        text = element_text(size = 10),
        panel.background = element_rect(fill = greyPalette5[1]),
        axis.title.x = element_text(size = small_size),
        axis.title.y.right = element_text(size = small_size),
        plot.subtitle = element_text(hjust = 0.5, size = base_size,  face="italic"),
        plot.title = element_text(hjust = 0.5, size = title_size),
        plot.caption=element_text(size = small_size, colour = "black", hjust = 0.5))
}


q5 <- function(variable) {as.factor(ntile(variable, 5))}

q <- function(variable) {as.factor(ntile(variable, 5))}

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                                  c(.01,.2,.4,.6,.8), na.rm=T),
                         digits = 3))
  }
}

qBr2 <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]]*100,0)/100,
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                                  c(.01,.2,.4,.6,.8), na.rm=T),
                         digits = 3))
  }
}

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <- get.knnx(measureTo, measureFrom, k)$nn.dist
  output <- as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  return(output)  
}


myCrossValidate <- function(dataset, id, dependentVariable, indVariables) {
  
  allPredictions <- data.frame()
  cvID_list <- unique(dataset[[id]])
  
  for (i in cvID_list) {
    
    thisFold <- i
    cat("This hold out fold is", thisFold, "\n")
    
    fold.train <- filter(dataset, dataset[[id]] != thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    fold.test  <- filter(dataset, dataset[[id]] == thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    
    regression <-
      glm(countMVTheft ~ ., family = "poisson", 
          data = fold.train %>% 
            dplyr::select(-geometry, -id))
    
    thisPrediction <- 
      mutate(fold.test, Prediction = predict(regression, fold.test, type = "response"))
    
    allPredictions <-
      rbind(allPredictions, thisPrediction)
    
  }
  return(st_sf(allPredictions))
}

#----------------------------------------------------------------------------------------------------------  

# Colors ("https://coolors.co/gradient-palette/a8f368-f9035e?number=7")
bluePalette5 <- c("#eff3ff","#bdd7e7","#6baed6","#3182bd","#08519c")
blue2Palette5 <- c("#08519c","#3182bd","#6baed6","#bdd7e7","#eff3ff")
orangePalette5 <- c("#FFF2E8","#FFD6B6","#FEB984","#FE9D51","#FD801F")
orange2Palette5 <- c("#FFDFD0","#FFB89F","#FF926E","#FF6B3D","#FF440C")
greyPalette5 <- c("#f7f7f7","#cccccc","#969696","#636363","#252525")
greenPalette5 <- c("#edf8e9","#bae4b3","#74c476","#31a354","#006d2c")
purplePalette5 <- c("#f2f0f7","#cbc9e2","#9e9ac8","#756bb1","#54278f")

#----------------------------------------------------------------------------------------------------------  

# LoadAPI(Min's key)
census_api_key("4bbe4bead4e5817f6a6b79e62c5bea69e77f1887", overwrite = TRUE)


```
</div>



## Ⅱ. Data

<div class = "blue">
```{r message=FALSE, warning=FALSE, include = FALSE, cache=TRUE}

# Tracts geometry, Census Information

policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271')

neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson")%>%
  st_transform('ESRI:102271')

#----------------------------------------------------------------------------------------------------------  

# update to 2018 data

# some of risk factors data hasn't been updated yet after 2018

Criminals <- read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2018/3i3m-jwuy")

#----------------------------------------------------------------------------------------------------------  

# check criminal distributions 

CriminalsType <- Criminals %>% dplyr::select(Primary.Type) %>% group_by(Primary.Type) %>% summarise(count = n())

CriminalsDescription <- Criminals %>% dplyr::select(Description) %>% group_by(Description) %>% summarise(count = n())

#----------------------------------------------------------------------------------------------------------  
  
# Dependent variable : MOTOR VEHICLE THEFT - AUTOMOBILE

MVTheft <- Criminals %>%
  filter(Primary.Type == "MOTOR VEHICLE THEFT" & Description == "AUTOMOBILE") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),Y = as.numeric(Y)) %>%
  dplyr::select(-Date, -Updated.On) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
  st_transform('ESRI:102271') %>%
  distinct()

#----------------------------------------------------------------------------------------------------------  

# Creating a fishnet grid

fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 500, 
               square = TRUE) %>%
  .[chicagoBoundary] %>%            # fast way to select intersecting polygons
  st_sf() %>%
  mutate(uniqueID = 1:n())

#----------------------------------------------------------------------------------------------------------  

# 6 risk factors

# risk factor 1 of 6 : Abandoned cars

rf1_abandonCars <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva")

rf1 <- rf1_abandonCars %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Cars")

# risk factor 2 of 6 : Abandoned Buildings

rf2_abandonBuildings <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/7nii-7srd") 

rf2 <- rf2_abandonBuildings %>%
    mutate(year = substr(date_service_request_was_received,1,4)) %>%  filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Buildings")

# risk factor 3 of 6 : graffiti

rf3_graffiti <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Graffiti-Removal-Historical/hec5-y4x5")

rf3 <- rf3_graffiti %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    filter(where_is_the_graffiti_located_ %in% c("Front", "Rear", "Side")) %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Graffiti")

# risk factor 4 of 6 : street Lights Out

rf4_streetLightsOut <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Street-Lights-All-Out/zuxi-7xem")

rf4 <- rf4_streetLightsOut %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Street_Lights_Out")
    
    
# risk factor 5 of 6 : sanitation

rf5_sanitation <- read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Sanitation-Code-Complaints-Hi/me59-5fac")

rf5 <- rf5_sanitation %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2018") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Sanitation")
    
    

# risk factor 6 of 6 : liquor Retail

rf6_liquorRetail <- read.socrata("https://data.cityofchicago.org/resource/nrmj-3kcf.json")

rf6 <- rf6_liquorRetail %>%
    filter(business_activity == "Retail Sales of Packaged Liquor") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Liquor_Retail")


```
</div>


## Ⅲ. Process

**1. Points Map of Automobile Theft**

description of what, when, and why you think selection bias may be an issue

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary, fill = greyPalette5[2], color = "transparent", alpha = 0.5) +
  geom_sf(data = MVTheft, colour=orangePalette5[4], size=0.01, show.legend = "point") +
  labs(title= "Automobile Theft",
       subtitle = "Total 8,800 cases",
       caption = "Figure 1-1") +
  mapThememin(),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = greyPalette5[2], color = "transparent", alpha = 0.5) +
  stat_density2d(data = data.frame(st_coordinates(MVTheft)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = orangePalette5[1], high = orangePalette5[5]) +
  scale_alpha(range = c(0.00, 0.2), guide = FALSE) +
  labs(title = "Density of Automobile Theft",
       subtitle = "2018 Chicago",
       caption = "Figure 1-2") +
  mapThememin() + theme(legend.position = "none"))

```
</div>


**2. Aggregate Automobile Theft to the Fishnet**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

crime_net <- 
  dplyr::select(MVTheft) %>% 
  mutate(countMVTheft = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countMVTheft = replace_na(countMVTheft, 0),
         uniqueID = 1:n(),
         cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE)) # cross-validation id approx 100 folds

ggplot() +
  geom_sf(data = crime_net, aes(fill = q5(countMVTheft)), color = "transparent")+
  scale_fill_manual(values = orangePalette5,
                    labels = qBr(crime_net, "countMVTheft"),
                    name = "Automobile\nTheft counts\n(quantiles)",
                    na.translate=FALSE) +
  labs(title = "Aggregate Map of Automobile Theft",
       subtitle = "2018 Chicago",
       caption = "Figure 2-1") +
  mapThememin()


```
</div>


**3. Multiple map of Automobile Theft risk factors in the fishnet**

counts, distance and/or other feature engineering approaches

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}


vars_net <- 
  rbind(rf1,rf2,rf3,rf4,rf5,rf6) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend, count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()


vars_net <- vars_net %>%
    mutate(Abandoned_Cars_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf1), k = 3)),
           Abandoned_Buildings_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf2), k = 3)),
           Graffiti_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf3), k = 3)),
           Street_Lights_Out_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf4), k = 3)),
           Sanitation_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf5), k = 3)),
           Liquor_Retail_nn3 = round(nn_function(st_c(st_coid(vars_net)), st_c(rf6), k = 3)))

vars_net.nn <- 
  dplyr::select(vars_net, ends_with("nn3")) %>%
    gather(Variable, value, -geometry)

vars <- unique(vars_net.nn$Variable)
mapList <- list()


for(i in vars){
  mapList[[i]] <-
    ggplot() +
    geom_sf(data = filter(vars_net.nn, Variable == i), aes(fill=q5(value)), color = "transparent") +
    scale_fill_manual(values = bluePalette5,
                      labels = qBr(vars_net.nn, "value"),
                      name = "Automobile\nTheft counts\n(quantiles)",
                      na.translate=FALSE) +
    labs(title = i) +
  mapThememin()
}

do.call(grid.arrange,c(mapList, ncol=3, top="Risk Factors by Fishnet"))

for(i in vars){
  mapList[[i]] <-
    ggplot() +
    geom_sf(data = filter(vars_net.nn, Variable == i), aes(fill=value), color = "transparent") +
    scale_fill_gradient(low = bluePalette5[1], high = bluePalette5[5]) +
    labs(title = i) +
  mapThememin()
}

do.call(grid.arrange,c(mapList, ncol=3, top="Risk Factors by Fishnet"))

#----------------------------------------------------------------------------------------------------------

```
</div>


**4. Local Moran’s I-related small multiple map of your outcome**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

#----------------------------------------------------------------------------------------------------------

final_net <-
  left_join(crime_net, st_drop_geometry(vars_net), by="uniqueID") 

final_net <-
  st_centroid(final_net) %>%
    st_join(dplyr::select(neighborhoods, name)) %>%
    st_join(dplyr::select(policeDistricts, District)) %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()

#----------------------------------------------------------------------------------------------------------

final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)

final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)

#----------------------------------------------------------------------------------------------------------

local_morans <- localmoran(final_net$countMVTheft, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

#----------------------------------------------------------------------------------------------------------

# join local Moran's I results to fishnet
final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(Count_Theft = countMVTheft, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)

#----------------------------------------------------------------------------------------------------------

vars2 <- unique(final_net.localMorans$Variable)
varList2 <- list()

for(i in vars2){
  varList2[[i]] <- 
    ggplot() +
    geom_sf(data = filter(final_net.localMorans, Variable == i), aes(fill=Value), color = "transparent") +
    scale_fill_gradient(low = orangePalette5[1], high = orangePalette5[5]) +
    labs(title = i) +
  mapThememin() + theme(legend.position="bottom")}


do.call(grid.arrange,c(varList2, ncol = 4, top = "Local Morans I statistics"))

#----------------------------------------------------------------------------------------------------------

final_net <-
  final_net %>% 
  mutate(theft.isSig = ifelse(local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(theft.isSig.dist = nn_function(st_c(st_coid(final_net)), st_c(st_coid(filter(final_net, theft.isSig == 1))), k = 1))

ggplot() +
  geom_sf(data = final_net, aes(fill=q5(theft.isSig.dist)), color = "transparent") +
  scale_fill_manual(values = orangePalette5,
                    labels = qBr(final_net, "theft.isSig.dist"),
                    name = "Distance(ft)\n(quantiles)",
                    na.translate=FALSE)+
  labs(title = "Distance from Automobile Theft Hotspot",
       subtitle = "Hotspot : p-value ≤ 0.001",
       caption = "Figure 4-5") +
  mapThememin()
  
```
</div>


**5. Multiple Scatterplots with correlations**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.height = '100%' }

correlation.long <-
  st_drop_geometry(final_net) %>%
    dplyr::select(-uniqueID, -cvID, -name, -District, -theft.isSig, -theft.isSig.dist) %>%
    gather(Variable, Value, -countMVTheft)

correlation.cor <-
  correlation.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, countMVTheft, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, countMVTheft)) +
  geom_point(size = 0.1, color = bluePalette5[3])  +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.2, size = 2, color = bluePalette5[5]) +
  geom_smooth(method = "lm", se = FALSE, colour = bluePalette5[5]) +
  facet_wrap(~Variable, ncol = 4, scales = "free") +
  labs(title = "Scatterplots of risk factors with correlations",
       subtitle = "Automobile theft counts as a function of risk factors",
       x = "risk factor value",
       y = "Automobile theft counts") +
  corTheme()

```
</div>


**6. A histogram of dependent variable**

Take a look at the skewed distribution of countBurglaries in the topmost histogram of Figure 5.15. Given burglary is a relatively rare event, it is reasonable for most grid cells to contain no crime events. When data is distributed this way, an OLS regression is inappropriate. In this section, a Poisson Regression is estimated which is uniquely suited to modeling a count outcome like countBurglaries.

There are many different approaches to modeling burglary counts. Here, a Poisson Regression is used, which is based on a Poisson distribution, simulated in the bottommost histogram of Figure 5.15. Does the observed and simulated distributions appear similar? There are many flavors of count-based regression, but the one used here is the most simple.43


<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

ggplot(final_net, aes(x = countMVTheft))+
  geom_histogram(fill=orangePalette5[3], color = "white", size = 1, xlim=c(20,50))+
  labs(title = "a Histogram of Dependent Variable",
       subtitle = "Automobilt theft in Chicago, 2018",
       x = "Automobilt theft",
       y = "counts",
       caption = "Figure 6-1")+
  geom_hline(yintercept = 0, size = 0.5, color = greyPalette5[2]) +
  corTheme2()+
  theme(panel.background = element_blank(),
        panel.grid.major.y = element_line(size = 0.3, color = greyPalette5[1]))  

```
</div>



**7. Maps of model errors**

by k-fold and spatial cross validation

<div class = "blue">
```{r message=FALSE, eval=TRUE, warning=FALSE, out.width = '100%'}

## Defined the variables we want, two types

reg.vars <- c("Abandoned_Buildings_nn3", "Abandoned_Cars_nn3", "Graffiti_nn3", 
              "Liquor_Retail_nn3", "Street_Lights_Out_nn3", "Sanitation_nn3")

reg.ss.vars <- c("Abandoned_Buildings_nn3", "Abandoned_Cars_nn3", "Graffiti_nn3", 
              "Liquor_Retail_nn3", "Street_Lights_Out_nn3", "Sanitation_nn3", "theft.isSig", "theft.isSig.dist")

#----------------------------------------------------------------------------------------------------------

# 100 fold regression

reg.cv <- myCrossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countMVTheft",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = cvID, countMVTheft, Prediction, geometry)

reg.ss.cv <- myCrossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countMVTheft",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = cvID, countMVTheft, Prediction, geometry)

#----------------------------------------------------------------------------------------------------------

# neighborhood fold regression
  
reg.spatialCV <- myCrossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countMVTheft",
  indVariables = reg.vars) %>%
    dplyr::select(cvID = name, countMVTheft, Prediction, geometry)

reg.ss.spatialCV <- myCrossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countMVTheft",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = name, countMVTheft, Prediction, geometry)

#----------------------------------------------------------------------------------------------------------

## Summary

reg.summary <- 
  rbind(
    mutate(reg.cv, 
           Error = Prediction - countMVTheft,
           Regression = "Random k-fold CV: Just Risk Factors"),
    mutate(reg.ss.cv,        
           Error = Prediction - countMVTheft,
           Regression = "Random k-fold CV: Spatial Process"),
    
    mutate(reg.spatialCV,    
           Error = Prediction - countMVTheft,
           Regression = "Spatial LOGO-CV: Just Risk Factors"),
                             
    mutate(reg.ss.spatialCV, 
           Error = Prediction - countMVTheft,
           Regression = "Spatial LOGO-CV: Spatial Process")) %>%
    st_sf() 

#----------------------------------------------------------------------------------------------------------


ggplot(reg.summary) +
  geom_sf(aes(fill = q5(Error)), color = "transparent" ) +
  scale_fill_manual(values = bluePalette5,
                    labels = qBr(reg.summary, "Error"),
                    name = "Error",
                    na.translate=FALSE) +
  labs(title = "Maps of Model Errors", 
       subtitle = "Errors = Prediction - Observation",
       caption = 'Figure 7-1') +
  facet_wrap(c(~Regression), ncol=4) +
  mapThememin()

#----------------------------------------------------------------------------------------------------------

error_by_reg_and_fold <- 
  reg.summary %>%
    group_by(Regression, cvID) %>% 
    summarize(Mean_Error = mean(Prediction - countMVTheft, na.rm = T),
              MAE = mean(abs(Mean_Error), na.rm = T),
              SD_MAE = mean(abs(Mean_Error), na.rm = T)) %>%
  ungroup()

error_by_reg_and_fold %>%
  ggplot(aes(MAE)) + 
  geom_histogram(bins = 30,  color = "white", size = 1, fill=bluePalette5[3]) +
  facet_wrap(~Regression, scales = "free_y") +  
  geom_hline(yintercept = 0, size = 0.5, color = greyPalette5[2]) +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) + 
  labs(title="Distribution of MAE", subtitle = "k-fold cross validation vs. LOGO-CV",
       x="Mean Absolute Error", y="Count") +
  corTheme2() +
  theme(panel.background = element_blank(),
        panel.grid.major.y = element_line(size = 0.3, color = greyPalette5[1]))


```
</div>


**8. A table of MAE and standard deviation of MAE by regression**

?
<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

st_drop_geometry(error_by_reg_and_fold) %>%
  group_by(Regression) %>% 
    summarize(Mean_MAE = round(mean(MAE), 2),
              SD_MAE = round(sd(MAE), 2)) %>%
  kable(caption = "<center><span style='font-size:12px; color: black; font-family: Arial;'>Table 8-1. MAE and standard deviation MAE by regression</span>", align = "c") %>%
    kable_minimal(full_width = T, html_font = "Arial", font_size = 14) %>%
    row_spec(2, color = greyPalette5[5], background = greyPalette5[1], bold = T) %>%
    row_spec(4, color = greyPalette5[5], background = greyPalette5[1], bold = T) %>%
    row_spec(0:4, extra_css = "line-height: 30px")
    

```
</div>


**9. A table of raw errors by race context**

for a random k-fold vs. spatial cross validation regression.

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

census_var2018 <- c("B25026_001E", "B02001_002E", "B19013_001E", "B25058_001E", "B06012_002E")
Chicago.tracts18 <- get_acs(geography = "tract", 
                           variables = census_var2018, 
                           year=2018, 
                           state="IL",
                           county= 031,
                           geometry=TRUE, 
                           output="wide") %>%
    dplyr::select(-geometry, -B25026_001M, -B02001_002M, -B19013_001M, -B25058_001M, -B06012_002M) %>% # Remove margin of error columns
    st_transform('ESRI:102271') %>%
  rename(TotalPop = B25026_001E,
         Whites = B02001_002E,
         MedHHInc = B19013_001E,
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2018") %>%
  mutate(pctWhite.scaled = pctWhite * 100,
         pctPoverty.scaled = pctPoverty * 100,
         raceContext = ifelse(pctWhite > .5, "Majority_White", "Majority_Non_White")) %>%
  dplyr::select(-Whites, -TotalPoverty, -NAME)

ggplot() +
  geom_sf(data = Chicago.tracts18, aes(fill = raceContext), color = greyPalette5[1], size = 0.1) +
  scale_fill_manual(values=c(bluePalette5[2], bluePalette5[4])) +
  labs(title = "Race context Map",
       subtitle = "Chicago, 2018",
       caption = "Figure 9-1") +
  mapThememin()


raceContext <-
  reg.summary %>% 
  st_centroid() %>%
  st_join(Chicago.tracts18) %>%
  na.omit() %>%
  st_drop_geometry() %>%
  group_by(Regression, raceContext) %>%
  summarize(mean.Error = mean(Error, na.rm = T)) %>%
  spread(raceContext, mean.Error) %>%
  kable(caption = "<center><span style='font-size:12px; color: black; font-family: Arial;'>Table 9-1. Mean Error by Neighborhood Racial Context</span>", align = "c") %>%
  kable_minimal(full_width = T, html_font = "Arial", font_size = 13) %>%
  row_spec(2, color = greyPalette5[5], background = greyPalette5[1], bold = T) %>%
  row_spec(4, color = greyPalette5[5], background = greyPalette5[1], bold = T) %>%
  row_spec(0:4, extra_css = "line-height: 30px")

raceContext
  
```
</div>


**10. The map comparing kernel density to risk predictions for the next year’s crime.**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

theft_ppp <- as.ppp(st_coordinates(MVTheft), W = st_bbox(final_net))
theft_KD.1000 <- spatstat.core::density.ppp(theft_ppp, 1000)

as.data.frame(theft_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
   ggplot() +
     geom_sf(aes(fill=value)) +
     geom_sf(data = sample_n(MVTheft, 1500), size = .5) +
     scale_fill_viridis(name = "Density") +
     labs(title = "Kernel density of 2018 MVTheft") +
     mapTheme()

## Getting 2019 theft data

MVTheft19 <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2019/3i3m-jwuy") %>% 
  filter(Primary.Type == "MOTOR VEHICLE THEFT" & Description == "AUTOMOBILE") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),
         Y = as.numeric(Y)) %>% 
  dplyr::select(-Date, -Updated.On) %>%
  na.omit %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:102271') %>% 
  distinct() %>%
  .[fishnet,]

theft_KDE_sum <- as.data.frame(theft_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) 

kde_breaks <- classIntervals(theft_KDE_sum$value, 
                             n = 5, "fisher")
theft_KDE_sf <- theft_KDE_sum %>%
  mutate(label = "Kernel Density",
         Risk_Category = ntile(value, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(MVTheft19) %>% mutate(theftCount = 1), ., sum) %>%
    mutate(theftCount = replace_na(theftCount, 0))) %>%
  dplyr::select(label, Risk_Category, theftCount)

theft_risk_sf <-
  reg.ss.spatialCV %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
         Risk_Category >= 90 ~ "90% to 100%",
         Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
         Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
         Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
         Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(MVTheft19) %>% mutate(theftCount = 1), ., sum) %>%
      mutate(theftCount = replace_na(theftCount, 0))) %>%
  dplyr::select(label,Risk_Category, theftCount)

```
</div>


<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

rbind(theft_KDE_sf, theft_risk_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    geom_sf(data = sample_n(MVTheft19, 3000), size = .5, colour = "black") +
    facet_wrap(~label, ) +
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Comparison of Kernel Density and Risk Predictions",
         subtitle="2018 theft risk predictions; 2019 theft") +
    mapTheme(title_size = 14)

```
</div>


**11. The bar plot making this comparison.**

<div class = "blue">
```{r message=FALSE, warning=FALSE, out.width = '100%'}

rbind(theft_KDE_sf, theft_risk_sf) %>%
  st_drop_geometry() %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countTheft = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Pcnt_of_test_set_crimes = countTheft / sum(countTheft)) %>%
    ggplot(aes(Risk_Category,Pcnt_of_test_set_crimes)) +
      geom_bar(aes(fill=label), position="dodge", stat="identity") +
      scale_fill_viridis(discrete = TRUE, name = "Model") +
      labs(title = "Risk prediction vs. Kernel density, 2018 theft",
           y = "% of Test Set Theft (per model)",
           x = "Risk Category") +
  theme_bw() +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

```
</div>


**12. Two paragraphs on why or why not you would recommend your algorithm be put into production..**



## Ⅳ. Results




























































